From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-17 dictionary.tex
QCS and LC for p-adic tori

\begin{corollary}
Let $T_K$ be an algebraic toris over $K$ and let $T$ be a Neron model for $T_K$.
The following is a commutative diagram of exact sequences,
after removing isomorphism groups.
\[
  \begin{tikzcd}[row sep=20, column sep=18]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \dar{\TrFrob{\pi_0(T)}}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d}{\TrFrob{\Gr_R(T)}} & \QCSiso{T^0} \arrow{r} \arrow{d}{\TrFrob{\Gr_R(T)^0}} & 0\\
    0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{infl'n}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{qc}(T(K),\EEx) \arrow{r}{\text{rest'n}} & \Hom_\text{qc}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\
     0 \arrow{r}  
 & \Hh^1(\Weil{}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{infl'n}}\arrow{d}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{rest'n}} \arrow{d}
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{}} \arrow{r}  \arrow{d}
 & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \]
Moreover, for $\cs{E}\in \QCS(\pi_0(T))$, $\cs{F}\in {\QCS(T)}$ and $\cs{F}^0\in {\QCS(T^0)}$, there are canonical isomorphisms
\[
\Aut_{\QCS(\pi_0(T))}(\cs{E})\iso (\check{T}_\ell)^{\Weil{K}},
\quad
\Aut_{\QCS(T)}(\cs{F})\iso (\check{T}_\ell)^{\Weil{K}},
\quad
\Aut_{\QCS(T^0)}(\cs{F}^0)\iso 1.
\]
 \end{corollary}
\begin{proof} 
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Accordingly, by Theorem~\ref{thm:CSXK}, the kernel of the trace of Frobenius is
\[
\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{}} = \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}}.
\] 
Also by Theorem~\ref{thm:CSXK}, $\Aut(\cs{F})  \iso \Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx)$, canonically.
By \cite{bitan:discriminant}*{Eq 3.1} again, 
\[
\Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx) \iso
\Hom(X_*(T_K)_{\Weil{K}},\EEx),
\]
canonically.
But $\Hom(X_*(T_K)_{\Weil{K}},\EEx) \iso \Hom(X_*(T_K),\EEx)^{\Weil{K}}$, canonically.
So, for any quasicharacter sheaf $\cs{F}$ for $T$,
\[
\Aut_\QCS{}(\cs{F}) \iso (\check{T}_\ell)^{\Weil{K}},
\]
canonically, where $\check{T}_\ell$ is the $\ell$-adic dual torus $\Hom(X_*(T_K),\EEx)$.
[under construction]
\end{proof}

>>>>

\begin{corollary}
Let $T_K$ be an algebraic toris over $K$ and let $T$ be a Neron model for $T_K$.
The following is a commutative diagram of exact sequences,
after removing isomorphism groups.
\[
  \begin{tikzcd}[row sep=20, column sep=18]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \dar{\TrFrob{\pi_0(T)}}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d}{\TrFrob{\Gr_R(T)}} & \QCSiso{T^0} \arrow{r} \arrow{d}{\TrFrob{\Gr_R(T)^0}} & 0\\
    0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{infl'n}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{qc}(T(K),\EEx) \arrow{r}{\text{rest'n}} & \Hom_\text{qc}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \]
 \end{corollary}
\begin{proof} 
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Accordingly, by Theorem~\ref{thm:CSXK}, the kernel of the trace of Frobenius is
\[
\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{}} = \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}}.
\] 
[under construction]
\end{proof}

\begin{corollary}
For $\cs{E}\in \QCS(\pi_0(T))$, $\cs{F}\in {\QCS(T)}$ and $\cs{F}^0\in {\QCS(T^0)}$, there are canonical isomorphisms
\[
\Aut_{\QCS(\pi_0(T))}(\cs{E})\iso (\check{T}_\ell)^{\Weil{K}},
\quad
\Aut_{\QCS(T)}(\cs{F})\iso (\check{T}_\ell)^{\Weil{K}},
\quad
\Aut_{\QCS(T^0)}(\cs{F}^0)\iso 1.
\]
\end{corollary}

\begin{proof}
By Theorem~\ref{thm:CSXK}, $\Aut(\cs{F})  \iso \Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx)$, canonically.
By \cite{bitan:discriminant}*{Eq 3.1} again, 
\[
\Hom((\pi_0(X)_{\bFq})_{\Weil{}},\EEx) \iso
\Hom(X_*(T_K)_{\Weil{K}},\EEx),
\]
canonically.
But $\Hom(X_*(T_K)_{\Weil{K}},\EEx) \iso \Hom(X_*(T_K),\EEx)^{\Weil{K}}$, canonically.
So, for any quasicharacter sheaf $\cs{F}$ for $T$,
\[
\Aut_\QCS{}(\cs{F}) \iso (\check{T}_\ell)^{\Weil{K}},
\]
canonically, where $\check{T}_\ell$ is the $\ell$-adic dual torus $\Hom(X_*(T_K),\EEx)$.
[under construction]
\end{proof}

\subsection{Geometric perspective on the Langlands correspondence for $p$-adic tori}

\[
\begin{tikzcd}
   0 \rar & \arrow[bend right=70]{dd} \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \dar{\TrFrob{\pi_0(T)}}
    & \QCSiso{T} \arrow{r}{\iota_0^*} \arrow[dotted, bend right=50]{dd}  \arrow{d}{\TrFrob{\Gr_R(T)}} & \QCSiso{T^0} \arrow[dotted, bend right=50]{dd}  \arrow{r} \arrow{d}{\TrFrob{\Gr_R(T)^0}} & 0\\
   0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{infl'n}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{qc}(T(K),\EEx) \arrow{r}{\text{rest'n}} & \Hom_\text{qc}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\    
    0 \arrow{r}  
 & \Hh^1(\Weil{}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{infl'n}}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{rest'n}} 
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{}} \arrow{r}  
 & 0
\end{tikzcd}
\]
